---
title: "Dissertation Mock up figures"
author: "Michelle DePrenger-Levin"
date: "November 7, 2019"
output: html_document
---

```{r}
library(raster)
library(gstat)
library(virtualspecies)
library(DiagrammeR)

```

# Spatial modeling: forecasting and descriptive    
gold standard    
when falls apart for rare plants    


<http://santiago.begueria.es/2010/10/generating-spatially-correlated-random-fields-with-r/>
```{r}
xy <- expand.grid(1:100, 1:100)
names(xy) <-  c('x','y')
# where formula defines the dependent variable (z) as a linear model of independent variables. For ordinary and simple kriging we can use the formula z~1; for simple kriging it is necessary to define a beta parameter too (see below); for universal kriging, if z is linearly dependent on x and y use the formula z~x+y. We are using simple kriging here.
# locations define the data coordinates, e.g. ~x+y in our case here. 
# dummy is a logical value, and it needs to be TRUE for unconditional simulation. 
# beta is used only for simple kriging, and is a vector with the trend coefficients (including an intercept); 
# if no independent variables are defined the model only contains an intercept, i.e. the simple kriging mean. model defines the variogram model, as defined by a call to vgm. vgm allows defining the (partial) sill, range and nugget paramaters, as well as the variogram model type (e.g. exponential, gaussian, spherical, etc). Anisotropy can also be used. 
# nmax defines the number of nearest observations that should be used for a kriging prediction or simulation.

g.dummy <- gstat(formula=z~1, locations=~x+y, dummy=T, beta=1, model=vgm(psill=0.025, range=5, model='Exp'), nmax=20)

yy <- predict(g.dummy, newdata=xy, nsim=4)
gridded(yy) = ~x+y
spplot(obj=yy[1])

spplot(obj=yy[2])

```


```{r}
plot(yy[1])

g.dummy2 <- gstat(formula=z~1, locations=~x+y, dummy=T, beta=3, model=vgm(psill=0.025, range=20, model='Exp'), nmax=20)

yy2 <- predict(g.dummy2, newdata=xy, nsim=4)
gridded(yy2) = ~x+y

g.dummy3 <- gstat(formula=z~1+x+y, locations=~x+y, dummy=T, beta=c(1,0.01,0.005), 
                  model=vgm(psill=0.025, range=50, model='Exp'), nmax=20)

yy3 <- predict(g.dummy3, newdata=xy, nsim=4)
gridded(yy3) = ~x+y

spplot(yy[1])
spplot(yy2[1])
spplot(yy3[1])
```


```{r}
# Different sample sizes
simsp1 <- data.frame(x = runif(10, 1, 99), y = runif(10, 1,99) )
simsp10 <- data.frame(x = runif(100, 1, 99), y = runif(100, 1,99) )



plot(yy[2])
points(simsp1, col = "black", pch = 20)


plot(yy3[3])
points(simsp1, col = "black", pch = 20)


# make points in relation to each layer? 
yy[1]@data$sim1
```








PVA temporal modeling demograpahics    
MatrixSplit: Divided means A broken into process-based submatrices U (survival), F (sexual reproduction), and C (clonal reproduction)
```{r}

library(popdemo)

load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/PVA/Paniw etal/patterns_temporal_autocorrelation-master/COMPADRE_v.4.0.0.RData")

# example from Wall et al. 2012 
table(compadre$metadata$SpeciesAccepted) # how many matrixes there are A
compadre$mat[compadre$metadata$SpeciesAccepted == "Astragalus michauxii"]
compadre$metadata[compadre$metadata$SpeciesAccepted == "Astragalus michauxii",]

```


Get the A matrices   
```{r}
AsMi_A <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Astragalus michauxii"], '[[', 1)


```























My attempts at creating a simulated landscape, just use above
```{r}

r4 <- r3 <-  r2 <- r1 <- raster(ncol=100, nrow=100) # looks like it's by row

# r1 will be the yes or no of presence points based on the 'environmental' layers r2 and r3
r2[] <- rnorm(ncell(r2), mean = 100, sd = 50) # no autocorrelation

# Add some autocorrelation
pickcell <- matrix(1:10000, nrow = 100, byrow = TRUE)
pickcell[4,5]  # 305
pickcell[3,5]
which(pickcell==104, arr.ind = TRUE)

pickcell[which(pickcell==104, arr.ind = TRUE)]

r3[] <- 1:10000
# r2[] <- calc(r2, function(x) 
#   if((which(pickcell==x, arr.ind = TRUE)[1]>10 & which(pickcell==x, arr.ind = TRUE)[2]>10) &
#      (which(pickcell==x, arr.ind = TRUE)[1]<90 & which(pickcell==x, arr.ind = TRUE)[2]<90)){
#     rnorm(1, mean = x, sd = )
#   } else {
#       rnorm(1, mean = matrix(r3@data@values, nrow = 100, byrow = TRUE), sd = 50)
#     }

data.frame(matrix(r2@data@values, nrow=100))
# 152.0787853

x <- 0
repeat {
  x <- x+1
  startval <- which(pickcell==sample(1:10000, 1), arr.ind = TRUE)
  r2[startval]
if(x=100){
  break
}
  }



# in relation to another measure
beta1 <- 1
int <- 100
r3[] <- rnorm(ncell(r3), mean = 10, sd = 20)
r4[] <- rlnorm(ncell(r4), meanlog = 0, sdlog = 1)

plot(r2)

r1[] <- rbinom(ncell(r1), ) 

```


Population viability modeling: forecasting and descriptive
gold standard
when falls apart for rare plants




Example using PVA for predicting impact of seed harvest





























Verision 1: pre 2019-12-01 December 

Model uncertainty
1. pick a threshold  
2. model the impact of spatial uncertanty across a landscape where there is varying environemntal differences and the scale of landscape variables is near reality vs. the spatial range of rare species
3. various levels of autocorrelated environmental variables

```{r}

GS <- runif(100, min=100, max = 50000)
B1 <- 10000 #intercept
B2 <- -0.6 #slope
Uncertainty <- (B1 + GS*B2)
UnJit <- Uncertainty*rnorm(100, mean = .5, sd = .15)

plot(GS,jitter(Uncertainty,amount = 0.0005), xlab="Geographic distribution", ylab= "Uncertainty")
plot(GS, (UnJit+abs(min(UnJit)))/max((UnJit+abs(min(UnJit)))), xlab="Geographic distribution", ylab= "Uncertainty")
```

Triangle plot for rarity
```{r}
library(ade4)
```


```{r}
#Environmental space is correlated to range
Environmental_Specificity <- jitter(2 + GS*0.91, amount = 100)

# local abundance low, more likely to miss if in clumps so if more environemtal habitat specificity, then shoudl be more lower samples, less detection (Dennett 2018)
Local_Abundance <- jitter(1 + ES*0.1, amount = 100)
Geographic_Range <- GS

raretaxa <- data.frame(Geographic_Range, Environmental_Specificity, Local_Abundance)

triangle.plot(raretaxa) # label = c("Geographic range","Habitat specificity","Local Abundance"))

```








